Incentivizing Exploration by Heterogeneous Users

### 2018-06-06 Discussion-Proof.tex
making notes cleaner
fix theorem 9, check david edits

\begin{rtheorem}{Theorem}{\ref{rst:discrete}}
Under Assumption~\ref{a:discrete}, the Discrete-Preference Algorithm has expected payment budget bounded above by 
\begin{align*}
\pfedit{
O \left(\ARMNUM^2  d \cdot (\MAXR + \Diam d \sigma)\cdot
    \max\bigg\{\frac{2}{p},
    \left( \frac{\sigma \Diam d}{\TieCutoff} \right)^{5/2}
    \bigg\}
\right)
}
\end{align*}
and expected regret bounded above by 
\begin{align}
O \left(\MAXR\cdot \frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\MAXR \ARMNUM^2 d^3 \TieDensity^2 \Diam^2 \sigma^2}{\MinProb^2}
  \right).  \nonumber
\end{align}
\end{rtheorem}

\begin{proof}
The proof of the expected payment bound is the same as the proof of Theorem~\ref{rst:budget}, except that we now define $h(\ell) := \max \left( \frac{2}{\MinProb},
\pfedit{
\left( \frac{24 \sigma \ell \Diam d}{\TieCutoff} \right)^{5/2},
}
\left( \frac{48 \sigma \ell \TieDensity \Diam d}{\MinProb} \right)^{5/2}
\right)
\pfedit{
= 
\max \left( \frac{2}{\MinProb},
\left( \frac{24 \sigma \ell \Diam d}{\TieCutoff} \right)^{5/2}
\right)
}
$.

To prove the bound on the expected regret, one can first prove tightened versions of Lemmas~\ref{lem:no-incentives} and \ref{lem:numP},
which replace the $\exp(2/\MinProb)$ term with simply $2/\MinProb$.
In return, the length of phase $s$ is now bounded only be $\ARMNUM s$ instead of $\ARMNUM \log(s)$.
Substituting these changes into the proof\pfedit{s} of Theorem\pfedit{s}~\ref{rst:regret} \pfedit{and \ref{rst:budget}}, we obtain the claimed bounds.

@@

@@

\begin{rtheorem}{Theorem}{\ref{rst:known-p}}
Under Assumption~\ref{a:known-p}, the Known-\MinProb Algorithm has an expected payment budget bounded above by 
    \bcedit{
\begin{align*}
    O\left(\ARMNUM^2 d\cdot(R+Dd\sigma)\cdot \max\bigg\{ \left(\frac{24\sigma \ell \Diam d}{\TieCutoff} \right)^{5/2},
        \left( \frac{48 \sigma \ell \TieDensity \Diam d}{\MinProb} \right)^{5/2}\bigg\}\right)
\end{align*}
}
and an expected regret bounded above by
\bcedit{
\begin{align*}
O \left(\frac{\MAXR \ARMNUM \TieDensity^3 \Diam^3 d^3 \sigma^3}{\MinProb^3}
  + \frac{\MAXR \ARMNUM^2 d^3 \TieDensity^2 \Diam^2 \sigma^2}{\MinProb^2}
  + \frac{\ARMNUM\Diam^2 d^2\sigma^2 \TieDensity(\log(T))^3}{\MinProb}\right).
\end{align*}
}
\end{rtheorem}

\begin{proof}
  The proof of the expected payment bound follows Lemma~\ref{rst:budget} except for defining
    \bcedit{
    $h(\ell) := \max\big\{ \left(\frac{24\sigma \ell \Diam d}{\TieCutoff} \right)^{5/2},
        \left( \frac{48 \sigma \ell \TieDensity \Diam d}{\MinProb} \right)^{5/2}\big\}$
    }
(without including the $\exp(2/\MinProb)$ term).

>>>>

\begin{rtheorem}{Theorem}{\ref{rst:discrete}}
Under Assumption~\ref{a:discrete}, the Discrete-Preference Algorithm has expected payment \pfdelete{budget} bounded above by 
$\pfedit{O \left(\ARMNUM^2 /p \right)}$
and expected regret bounded above by \pfedit{$O(\ARMNUM / p)$}.
\end{rtheorem}

\begin{proof}
The proof of the expected payment bound is the same as the proof of Theorem~\ref{rst:budget}, except that we now define $h(\ell) := \max \left( \frac{2}{\MinProb},
\pfedit{
\left( \frac{24 \sigma \ell \Diam d}{\TieCutoff} \right)^{5/2}
}
\right).
$
%\pfedit{and $s_1 = \max(2,\frac{30\sigma^3}{x^3}, \frac2{\MinProb})$}.

To prove the bound on the expected regret, one can first prove tightened versions of Lemmas~\ref{lem:no-incentives} and \ref{lem:numP},
which replace the $\exp(2/\MinProb)$ term with \pfdelete{simply} $2/\MinProb$ \pfedit{and use $L=0$.}
In return, the length of phase $s$ is now bounded only by $\ARMNUM s$ instead of $\ARMNUM \log(s)$, \pfedit{and the expectetd number of times steps in which a payment is made is bounded by $O(N/p + N^2)$.}
Substituting these changes into the proof of Theorem~\ref{rst:regret}, we obtain the bound $O(R(N/p + N^2) + NR + NR) = O(N/p)$ since $p \le 1/N$.

@@

\pfcomment{
The calculation I did for the bound on the regret is:

\begin{itemize}
\item regret from times when we pay is $RN(N + 1/p)$
\item regret from phases with $\gamma(s) >= \hat{z}$ is $O(NR D^6 d^6 \sigma^6 / \hat{z}^6)$
\item regret from pulls with small regret is 0
\item regret from pulls with big regret is $O(NR)$
\end{itemize}
}

@@

\begin{rtheorem}{Theorem}{\ref{rst:known-p}}
Under Assumption~\ref{a:known-p}, the Known-\MinProb Algorithm has an expected payment budget bounded above by 
    \bcedit{
\begin{align*}
    O\left(\ARMNUM^2 d\cdot(R+Dd\sigma)\cdot \max\bigg\{ \left(\frac{\sigma \Diam d}{\TieCutoff} \right)^{5/2},
        \left( \frac{\sigma \TieDensity \Diam d}{\MinProb} \right)^{5/2}\bigg\}\right)
\end{align*}
}
and an expected regret bounded above by
\bcedit{
\begin{align*}
O \left(\frac{\MAXR \ARMNUM \TieDensity^3 \Diam^3 d^3 \sigma^3}{\MinProb^3}
  + \frac{\MAXR \ARMNUM^2 d^3 \TieDensity^2 \Diam^2 \sigma^2}{\MinProb^2}
  + \frac{\ARMNUM\Diam^2 d^2\sigma^2 \TieDensity(\log(T))^3}{\MinProb}\right).
\end{align*}
}
\end{rtheorem}

\begin{proof}
  The proof of the expected payment bound follows Lemma~\ref{rst:budget} except for defining
    \bcedit{
    $h(\ell) := \max\big\{ \left(\frac{24\sigma \ell\Diam d}{\TieCutoff} \right)^{5/2},
        \left( \frac{48\sigma \ell\TieDensity \Diam d}{\MinProb} \right)^{5/2}\big\}$
    }
(without including the $\exp(2/\MinProb)$ term).

### 2018-06-06 Discussion.tex
changed the regret for Theorem 10
fixing theorem 9
addressed the issue in Theorem 10
my late nite attempt to fix thm 9
fix theorem 9, check david edits

\begin{assumption}[Discrete Preferences]
\label{a:discrete}
There is a positive $\delta_0$ such that
$\AlmostTied{\delta_0} = 0$.
\end{assumption}

@@

\begin{theorem}
\label{rst:discrete}
Under Assumption~\ref{a:discrete}, the Discrete-Preference Algorithm has expected payment \pfdelete{budget} bounded above by 
\begin{align*}
O \left(\frac{\ARMNUM^2 d \cdot (\MAXR + \Diam d \sigma)\sigma^{5/2}L^{5/2}\Diam^{5/2}d^{5/2}}{\MinProb^{5/2}} \right),
\end{align*}
and expected regret bounded above by 
\begin{align}
O \left(\MAXR\cdot \frac{\ARMNUM \TieDensity^4 \Diam^4 d^4 \sigma^4}{\MinProb^4}
  + \frac{\MAXR \ARMNUM^2 d^3 \TieDensity^2 \Diam^2 \sigma^2}{\MinProb^2}
  \right).  \nonumber
\end{align}
\end{theorem}

@@

\begin{theorem}
\label{rst:known-p}
Under Assumption~\ref{a:known-p}, the Known-\MinProb Algorithm has expected payment \pfdelete{budget} bounded above by 
\begin{align*}
O \left(\frac{\ARMNUM^2 d \cdot (\MAXR + \Diam d \sigma)\sigma^{5/2}L^{5/2}\Diam^{5/2}d^{5/2}}{\MinProb^{5/2}} \right),
\end{align*}
and expected regret bounded above by
\begin{align*}
O \left(\frac{\MAXR \ARMNUM \TieDensity^3 \Diam^3 d^3 \sigma^3}{\MinProb^3}
  + \frac{\MAXR \ARMNUM^2 d^3 \TieDensity^2 \Diam^2 \sigma^2}{\MinProb^2}
  + \frac{\ARMNUM^2 d\MAXR}{\MinProb}
  + \frac{\ARMNUM\Diam^2 d^2\sigma^2 \TieDensity(\log(T))^3}{\MinProb}\right).
\end{align*}
\end{theorem}

>>>>

\begin{assumption}[Discrete Preferences]
\label{a:discrete}
There is a positive \pfedit{$\hat{z}$} such that
$\AlmostTied{\pfedit{\hat{z}}} = 0$.
\end{assumption}

@@

\begin{theorem}
\label{rst:discrete}
Under Assumption~\ref{a:discrete}, the Discrete-Preference Algorithm has expected payment \pfdelete{budget} bounded above by 
$\pfedit{O \left(\ARMNUM^2 /p \right)}$
and expected regret bounded above by \pfedit{$O(\ARMNUM / p)$}.
\end{theorem}

@@

\begin{theorem}
\label{rst:known-p}
Under Assumption~\ref{a:known-p}, the Known-\MinProb Algorithm has expected payment \pfdelete{budget} bounded above by 
    \bcedit{
\begin{align*}
    O\left(\ARMNUM^2d\cdot(R+\Diam d\sigma)\cdot\max\bigg\{\left(\frac{\sigma \Diam d}{\TieCutoff}\right)^{5/2},
    \left(\frac{\sigma \TieDensity\Diam d}{\MinProb}\right)^{5/2}\bigg\}\right)
\end{align*}
}
and expected regret bounded above by
\bcedit{
\begin{align*}
O \left(\frac{\MAXR \ARMNUM \TieDensity^3 \Diam^3 d^3 \sigma^3}{\MinProb^3}
  + \frac{\MAXR \ARMNUM^2 d^3 \TieDensity^2 \Diam^2 \sigma^2}{\MinProb^2}
  + \frac{\ARMNUM\Diam^2 d^2\sigma^2 \TieDensity(\log(T))^3}{\MinProb}\right).
\end{align*}
}
\end{theorem}

### 2018-06-06 algorithm_regret.tex
fix theorem 9, check david edits

We distinguish between agents incurring large regret
(which requires severe misestimates of arm locations;
such misestimates are exponentially unlikely to occur), 
and agents incurring small positive regret,
which requires these agents to be almost tied in their preference for
the best arm.
To be more precise, we define (with foresight) a phase-dependent
cutoff
$\EarlyCut{s} = \sqrt{\frac{128\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}}$,
and consider a regret exceeding \EarlyCut{s} large.

For most of the proof, we will focus on the case $\EarlyCut{s} \leq \TieCutoff$.
The remaining phases are the ones with
$s \leq \frac{128}{1.8} \cdot \log(T) \cdot \Diam^2 d^2 \sigma^2/\TieCutoff^2$.
Each such phase lasts for at most $\ARMNUM \cdot \log(s)$ steps
in expectation by Lemma~\ref{lem:phase-length},
and each step incurs regret at most \MAXR.
The total expected regret incurred in these phases is at most
$O(\ARMNUM \MAXR \cdot \log(T) \Diam^2 d^2 \sigma^2/\TieCutoff^2
\cdot \log (\log(T) \cdot \Diam^2 d^2 \sigma^2/\TieCutoff^2)
\leq O(\ARMNUM \MAXR \log^2(T) \cdot \Diam^3 d^3 \sigma^3/\TieCutoff^3)$,
by crudely bounding $\log y \leq O(\sqrt{y})$.

@@

To bound the probability of \AccEU{\tau_s^k}{\frac{\EarlyCut{s}}{2\Diam d}},
we use Lemma~\ref{lem:round-prob}
(with $x_s = x'_s = \frac{\EarlyCut{s}}{2 \Diam d}$
and a stopping time of $\tau_s^k$).
We first verify that this choice of $x_s, x'_s$
satisfies the assumption of Lemma~\ref{lem:round-prob} for all
$s \geq 2$,
i.e., that
$\sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{n x_n^2}{16 \sigma^2}}
\leq \frac{n x_n}{2 \sigma}$
for all $n \geq 2$.
Substituting that
$x_n = \frac{\EarlyCut{n}}{2 \Diam d} = \sqrt{\frac{32\log(T) \cdot \sigma^2}{1.8 n}}$,
the left-hand side is

\[
  \sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{2 \log T}{1.8}}
  \; \leq \;
  \sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{2 n \log T}{1.8}},
\]
while the right-hand side is $\sqrt{\frac{8 n \log(T)}{1.8}}$.
Squaring both sides and canceling out common terms,
the desired inequality is equivalent to
$\frac{50}{9} \log T \geq \log(\log_{1.1}(n) + 1)$.
Because $n \leq T$, it is sufficient to show that
$\frac{50}{9} \log T \geq \log(\log_{1.1}(T) + 1)$,
which can be verified by numerical calculation for $T=2$ and a
derivative test.

Now, by Lemma~\ref{lem:round-prob} and a union bound over all arms $i$
and attributes $j$, we get that 
\begin{align*}
\Prob{\AccEU{\tau_s^k}{x_s}}
& \leq 24 \ARMNUM d \cdot \exp \left( \frac{-1.8 s x^2}{16 \sigma^2} \right)
\; = \; \frac{24 \ARMNUM d}{T^2}.
\end{align*}

In the low-probability event, we simply 
bound the maximum regret by \MAXR.
Summing over the $T$ time periods,
the total expected regret from large-regret steps is at most
$24 \ARMNUM d \MAXR / T$.

Adding all four types of regret terms,
and substituting 
$\EarlyCut[2]{s} = \frac{128\log(T) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$,
the cumulative regret up to time $T$ is at most
$
O \left(\ARMNUM \MAXR \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \MAXR \log^2(T)
+ \frac{\ARMNUM \MAXR d}{T}
+ \ARMNUM \TieDensity \cdot \sum_{s=1}^{T} \EarlyCut[2]{s} \log(s) \right)\\
= 
O \left(\ARMNUM  \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \log^2(T) + \ARMNUM \TieDensity \cdot \log^3(T) \right)$.\QED
\end{emptyextraproof}

>>>>

We distinguish between agents incurring large regret
(which requires severe misestimates of arm locations;
such misestimates are exponentially unlikely to occur), 
and agents incurring small positive regret,
which requires these agents to be almost tied in their preference for
the best arm.
To be more precise, we define (with foresight) a phase-dependent
cutoff
$\EarlyCut{s} = \sqrt{\frac{128\log(s) \cdot \Diam^2 d^2 \sigma^2}{1.8 s}}$,
and consider a regret exceeding \EarlyCut{s} large.

For most of the proof, we will focus on the case $\EarlyCut{s} \leq \TieCutoff$.
The remaining phases are the ones with
$\frac{s}{\log s} \leq \frac{128}{1.8} \cdot \Diam^2 d^2 \sigma^2/\TieCutoff^2$.
Because $\frac{s}{\log s} = \Omega(s^{2/3})$, there are at most
$O(\Diam^3 d^3 \sigma^3/\TieCutoff^3)$ such phases,
each such phase lasts for at most $\ARMNUM \cdot \log(s)$ steps
in expectation by Lemma~\ref{lem:phase-length},
and each step incurs regret at most \MAXR.
Therefore, the total expected regret incurred in these phases is at most
$O(\ARMNUM \MAXR \cdot \Diam^3 d^3 \sigma^3/\TieCutoff^3
\cdot \log (\Diam^3 d^3 \sigma^3/\TieCutoff^3))
\leq O(\ARMNUM \MAXR \Diam^4 d^4 \sigma^4/\TieCutoff^4)
= O(\ARMNUM \MAXR)$,
by crudely bounding $\log y \leq O(y^{1/3})$.

@@

To bound the probability of \AccEU{\tau_s^k}{\frac{\EarlyCut{s}}{2\Diam d}},
we use Lemma~\ref{lem:round-prob}
(with $x_s = x'_s = \frac{\EarlyCut{s}}{2 \Diam d}$
and a stopping time of $\tau_s^k$).
We first verify that this choice of $x_s, x'_s$
satisfies the assumption of Lemma~\ref{lem:round-prob} for all
$s \geq 2$,
i.e., that
$\sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{n x_n^2}{16 \sigma^2}}
\leq \frac{n x_n}{2 \sigma}$
for all $n \geq 2$.
Substituting that
$x_n = \frac{\EarlyCut{n}}{2 \Diam d} = \sqrt{\frac{32\log(n) \cdot \sigma^2}{1.8 n}}$,
the left-hand side is

\[
  \sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{2 \log n}{1.8}}
  \; \leq \;
  \sqrt{0.6 n \cdot \log (\log_{1.1}(n) + 1) + \frac{2 n \log n}{1.8}},
\]
while the right-hand side is $\sqrt{\frac{8 n \log(n)}{1.8}}$.
Squaring both sides and canceling out common terms,
the desired inequality is equivalent to
$\frac{50}{9} \log n \geq \log(\log_{1.1}(n) + 1)$.
This can be verified by numerical calculation for $n=2$ and a
derivative test.

Now, by Lemma~\ref{lem:round-prob} and a union bound over all arms $i$
and attributes $j$, we get that 
\begin{align*}
\Prob{\AccEU{\tau_s^k}{x_s}}
& \leq 24 \ARMNUM d \cdot \exp \left( \frac{-1.8 s x_s^2}{16 \sigma^2} \right)
\; = \; \frac{24 \ARMNUM d}{s^2}.
\end{align*}

In the low-probability event, we simply bound the maximum regret by \MAXR.
Summing over \dkedit{all} time periods,
the total expected regret from large-regret steps is at most
$24 \ARMNUM d \MAXR \cdot \frac{\pi^2}{6} = O(\ARMNUM \MAXR)$.

Adding all four types of regret terms,
and substituting 
$\EarlyCut[2]{s} = \frac{128 \dkedit{\log(s)} \cdot \Diam^2 d^2 \sigma^2}{1.8 s}$,
the cumulative regret up to time $T$ is at most
$
O \left(\ARMNUM \MAXR \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \MAXR
+ \ARMNUM \MAXR
+ \ARMNUM \TieDensity \cdot \sum_{s=1}^{T} \EarlyCut[2]{s} \log(s) \right)\\
= 
O \left(\ARMNUM  \cdot \exp \left( \frac{2}{p} \right)
+ \ARMNUM \TieDensity \cdot \log^3(T) \right)$.\QED
\end{emptyextraproof}
